---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# clueRIO <a href='https://github.com/luckinet/cluerio/'><img src='man/figures/logo.svg' align="right" height="200" /></a>

<!-- badges: start -->
[![CRAN status](https://www.r-pkg.org/badges/version/cluerio)](https://CRAN.R-project.org/package=cluerio)

[![R-CMD-check](https://github.com/luckinet/cluerio/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/luckinet/cluerio/actions/workflows/R-CMD-check.yaml)
[![Codecov test coverage](https://codecov.io/gh/luckinet/cluerio/branch/main/graph/badge.svg)](https://app.codecov.io/gh/luckinet/cluerio?branch=main)
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
<!-- badges: end -->

## Overview

The [CLUMondo](https://www.environmentalgeography.nl/site/data-models/models/clumondo-model/) model is a dynamic, spatially explicit, land use and land cover change model. clueRIO manages input/output for setting up the file-structure required for successfully configuring and running a CLUE model. This acts as an interface that enables users to build and document scenarios based on such models quickly and in a script-based fashion.

The most up-to-date version is typically a command line version. This R-package has the aim to simplify specifying all files needed to successfully run that version. Please make sure to reference not only this R-package, but also the references on the website and in particular *Van Asselen, S., Verburg, P.H., 2013. Land cover change or land-use intensification: simulating land system change with a global-scale land change model. Global Change Biology 19, 3648–3667*.

## Installation

Install the official version from CRAN:

```{r, eval=FALSE}
# install.packages("clueRIO")
```

Install the latest development version from github:

```{r, eval=FALSE}
devtools::install_github("luckinet/clueRIO")
```


## Examples

```{r, echo=FALSE, eval=FALSE}
# define metadata ----
# dir <- tempdir()
# data_dir <- paste0(system.file("test_datasets", package = "clueRIO", mustWork = TRUE), "/")

# lus <- c("forest", "cropland")
# commodities <- c("crops", "trees")
# targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"

# start project ----
# startProject(root = dir, name = "2lus_1demand_minimal", start = 2000, end = 2019,
#              landuse = lus, demands = commodities)

# register gridded objects ----
# LUS <- tibble(system = lus) %>%
#   mutate(ID = seq_along(system)-1)

# regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
# setLAT(table = LUS)
# regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
# regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest_year0", crs = targetCRS)

# register landuse systems ----
# regLanduse(name = "forest", resistance = 0.5) %>%
#   setConversion(to = "cropland", label = "intensification") %>%
#   setSuitability(const = 1, forest_suit = 1)
# regLanduse(name = "cropland", resistance = 0.5) %>%
#   setSuitability(const = 1, cropland_suit = 1)

# register demand types ----
# regCommodity(name = "crops") %>%
#   setProduction(by = "cropland", amount = 1)

# register regions ----
# st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#   mutate(region = "somewhere") %>%
#   regRegions(region = "region")

# register demand ----
# tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#        grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#   regDemand(values = "grown", region = "region")

# check out the landuse change sequence ----
showSequence()

# prepare input files ----
makeModel()
setOptions(out_write = 2)

# and finally run the model ----
runModel()

# make an animation ----
showAnimation(interval = 0.5)
```


```{r, eval=FALSE}
library(clueRIO)

data_dir <- system.file("test_datasets", package = "clueRIO", mustWork = TRUE)

regions <- st_read(dsn = paste0(data_dir, "/example_region.gpkg")) |> 
  mutate(region = "somewhere")

# landSystems <- tibble(ID = c(), system = c())

minimal <- clue_scene(
  name = "2lus_1demand_minimal", 
  period = c(2000, 2019),
  root = tempdir(), 
  description = "this is a minimal showcase where two landuse ...")

# first, specify the land system basis ...
minimal <- minimal |> 
  clue_landsystems(file = paste0(data_dir, "/landuse.tif"), 
                   regions = regions, 
                   cats = landSystems)

# ... then the drivers used to determine patterns
minimal <- minimal |> 
  clue_driver(file = paste0(data_dir, "/cropland_suit.tif"), 
              name = "suitCrop") |> 
  clue_driver(file = paste0(data_dir, "/forest_suit.tif"), 
              name = "suitFor")

# set land types up
minimal <- minimal |> 
  clue_landtype(name = "forest", resistance = 0.5,
                suitability = tibble(driver = "suitFor", 
                                     coef = 1)) |> 
  clue_landtype(name = "cropland", resistance = 0.5,
                suitability = tibble(driver = "suitCrop", 
                                     coef = 1), 
                production = tibble(commodity = "crops", 
                                    amount = 1))

# set commodities up
minimal <- minimal |> 
  clue_commodity(name = "crops", 
                 demand = tibble(year = c(2000:2019), 
                                 region = "somewhere", 
                                 amount = seq(from = 4760, to = 5000, length.out = 20)))

minimal <- minimal |> 
  clue_conversion(from = "forest", to = "cropland", label = "intensification")

# possibly adapt options
# clue_options(...)

# initiate the model
clue_initiate(scene = minimal)

# possibly check whether everything is set up properly
# clue_validate(scene = minimal)

# and finally run the model
clue_run(scene = minimal)
```


