---
title: "Getting Started"
format:
  html:
    toc: true
    theme: 
      light: flatly
      dark: darkly
    html-math-method: katex
    number-sections: true
    anchor-sections: true
    code-tools: true
    code-copy: true
    code-link: true
    code-fold: true
    code-summary: "Show the code"
comments:
  giscus: 
    repo: luckinet/clueRIO
highlight-style: github
fontsize: 12pt
linestretch: 1.15
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{getting_started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
#| eval: false

library(clueRIO)
library(tidyverse)
```

To get you started on building CLUMondo models with the help of `clueRIO`, we go through a little tutorial of the essential steps required to build, parameterize and run such a model.

```{r}
#| eval: false

data <- system.file("test_data", package = "clueRIO", mustWork = TRUE)

# our example model will be called '2 land systems (ls2), 1 demand (d1)'.
model <- "ls2d1"

# open a new scene for the model
minimal <- clue_scene(
  root = paste0(tempdir(), "/", model), 
  name = model, 
  period = c(2000, 2024),
  regions = paste0(data, "/test_region.gpkg"),
  description = "this is a minimal showcase where forest is converted to cropland to produce crops.")

# first, specify the gridded items ...
# ... initial land systems
minimal <- minimal |> 
  clue_gridded(file = paste0(data, "/ls2.tif"),
               type = "initial")

# ... then suitability drivers
minimal <- minimal |> 
  clue_gridded(file = paste0(data, "/suitability1.tif"), 
               type = "driver", name = "cropland_suit") |> 
  clue_gridded(file = paste0(data, "/suitability2_0.tif"),
               type = "driver", name = "forest_suit")

# setup land types
minimal <- minimal |> 
  # forest shall be convertable to cropland
  clue_landtype(name = "forest", resistance = 0.5,
                suitability = tibble(driver = c("const", "forest_suit"),
                                     beta = c(1, 1)),
                conversion = tibble(to = "cropland", 
                                    label = "intensification")) |> 
  # cropland shall produce crops
  clue_landtype(name = "cropland", resistance = 0.5,
                suitability = tibble(driver = c("const", "cropland_suit"),
                                     beta = c(1, 1)),
                production = tibble(name = "crops", 
                                    amount = 1, priority = 1))


# setup goods and services
minimal <- minimal |> 
  clue_goods(name = "crops", 
             demand = tibble(year = c(2000:2024), 
                             region = "ravenholm", 
                             amount = seq(from = 4760, to = 5000, length.out = 25)))

# possibly adapt options
# clue_options(...)

# initiate the model
clue_initiate(scene = minimal)

# and finally run the model
clue_run(scene = minimal, clue.exe = ...)

# possibly backup the model outputs
# clue_backup(scene = minimal)
```

Instead of specifying items step by step, one can also streamline this. The functions in this package are optimized to support batch-processing or quickly setting up scenarios where dimensions can be varied according to some scheme. The following shows a possible workflow.

```{r}
#| eval: false

# open a new scene for the model, like before
model <- "ls2d1"
minimal <- clue_scene(
  root = paste0(tempdir(), "/", model), 
  name = model, 
  period = c(2000, 2024),
  regions = paste0(data, "/test_region.gpkg"),
  description = "this is a minimal showcase where forest is converted to cropland to produce crops.")

# the following items would be produced in some external script(s) ...
.grid <- tibble(file = paste0(data, c("/ls2.tif", "/suitability1.tif", "/suitability2_0.tif")), 
                type = c("initial", "driver", "driver"),
                name = c(NA_character_, "cropland_suit", "forest_suit"))
.suit <- tibble(type = c("forest", "forest", "cropland", "cropland"),
                driver = c("const", "forest_suit", "const", "cropland_suit"),
                beta = c(1, 1, 1, 1))
.conv <- tibble(type = "forest",
                to = "cropland",
                label = "intensification")
.prod <- tibble(type = "cropland", 
                name = "crops",
                amount = 1,
                priority = 1)
.demd <- tibble(name = "crops", 
                year = c(2000:2024), 
                region = "ravenholm", 
                amount = seq(from = 4760, to = 5000, length.out = 25))

#... and can then be easily fed into the functions
# for example with a for loop (if there are too many items for typing them all)
for(grd in seq_along(.grid)){
  
  minimal <- .grid |> 
    filter(row_number() == grd) |> 
    {\(x) clue_gridded(scene = minimal, file = x$file, type = x$type, name = x$name)}()
  
}

# or more classically (even though this could be further streamlined just as above)
minimal <- minimal |>
  clue_landtype(name = "forest", resistance = 0.5,
                suitability = .suit |> filter(type == "forest") |> select(-type),
                conversion = .conv |> filter(type == "forest") |> select(-type)) |>
  clue_landtype(name = "cropland", resistance = 0.5,
                suitability = .suit |> filter(type == "cropland") |> select(-type),
                production = .prod |> filter(type == "cropland") |> select(-type),) |> 
  clue_goods(name = "crops", demand = .demd |> filter(name == "crops") |> select(-name)) |> 
  clue_initiate()

clue_run(scene = minimal, clue.exe = ...)
```

