---
title: "cluRIO use case overview"
format:
  html:
    toc: true
    theme: 
      light: flatly
      dark: darkly
    html-math-method: katex
    number-sections: true
    anchor-sections: true
    code-tools: true
    code-copy: true
    code-link: true
    code-fold: true
    code-summary: "Show the code"
comments:
  giscus: 
    repo: luckinet/clueRIO
highlight-style: github
fontsize: 12pt
linestretch: 1.15
urlcolor: blue
vignette: >
  %\VignetteIndexEntry{use_cases}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

This is a (incomplete) list of potential use cases of this software. They all are reflected by unit tests included in this package and show what functions are required at minimum to build a valid [CLUMondo](https://www.environmentalgeography.nl/site/data-models/models/clumondo-model/) model.


## Example data

Let's first load the requried packages and define the path for example data.

```{r setup}
library(clueRIO)
library(tibble)

data <- system.file("test_data", package = "clueRIO", mustWork = TRUE)
```

There are various tables and raster layers used throughout this tutorial.

```{r}
# plot the rasters
```


# minimal example

In a system with cropland and forest, convert forest to cropland at a rate required only by the demand of crops.

```{r}
#| eval: false

model <- "ls2d1"

minimal <- clue_scene(
  root = paste0(tempdir(), "/", model), 
  name = model, 
  period = c(2000, 2024),
  regions = paste0(data, "/test_region.gpkg"),
  description = "this is a minimal showcase where forest is converted to cropland to produce crops.") |> 
  clue_gridded(file = paste0(data, "/ls2.tif"),
               type = "initial") |> 
  clue_gridded(file = paste0(data, "/suitability1.tif"), 
               type = "driver", 
               name = "cropland_suit") |> 
  clue_gridded(file = paste0(data, "/suitability2_0.tif"),
               type = "driver", 
               name = "forest_suit") |> 
  clue_landtype(name = "forest", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "forest_suit"),
                                     beta = c(1, 1)),
                conversion = tibble(to = "cropland", 
                                    label = "intensification")) |> 
  clue_landtype(name = "cropland", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "cropland_suit"),
                                     beta = c(1, 1)),
                production = tibble(goods = "crops", 
                                    amount = 1, priority = 1)) |> 
  clue_goods(name = "crops", 
             demand = tibble(year = c(2000:2024), 
                             region = "ravenholm", 
                             amount = seq(from = 4760, to = 5000, length.out = 25))) |> 
  clue_initiate()

clue_run(scene = minimal, clue.exe = ...)
```

## minimal example, stepwise

In a system with cropland and forest, convert forest to cropland at a rate required only by the demand of crops. Here, each year is calculated step by step (allowing further modules to interact with the input data of each new time step).

```{r}
#| eval: false

model <- "ls2d1_step"

minimal <- clue_scene(
  root = paste0(tempdir(), "/", model), 
  name = model, 
  period = c(2000, 2024),
  regions = paste0(data, "/test_region.gpkg"),
  description = "this is a minimal showcase where forest is converted to cropland to produce crops.") |> 
  clue_gridded(file = paste0(data, "/ls2.tif"),
               type = "initial") |> 
  clue_gridded(file = paste0(data, "/suitability1.tif"), 
               type = "driver", 
               name = "cropland_suit") |> 
  clue_gridded(file = paste0(data, "/suitability2_0.tif"),
               type = "driver", 
               name = "forest_suit") |> 
  clue_landtype(name = "forest", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "forest_suit"),
                                     beta = c(1, 1)),
                conversion = tibble(to = "cropland", 
                                    label = "intensification")) |> 
  clue_landtype(name = "cropland", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "cropland_suit"),
                                     beta = c(1, 1)),
                production = tibble(goods = "crops", 
                                    amount = 1, priority = 1)) |> 
  clue_goods(name = "crops", 
             demand = tibble(year = c(2000:2024), 
                             region = "ravenholm", 
                             amount = seq(from = 4760, to = 5000, length.out = 25))) |> 
  clue_initiate(module = "stepwise")

clue_run(scene = minimal, clue.exe = ...)
```

## minimal example, with mask

There are various ways in which a mask can be employed.

::: {.panel-tabset}
## protected areas

...

```{r}
#| eval: false

model <- "ls2d1_protected"

```

```{r}
#| eval: false
#| echo: false

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing/"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("forest", "cropland")
#   commodities <- c("crops", "trees")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "wuf_protected", start = 2000, end = 2019,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   LUS <- tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1)
#
#   regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
#   setLAT(table = LUS)
#   regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
#   regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest", crs = targetCRS)
#   regGridded(name = "region", dir = data_dir, pattern = "protected", crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "forest", resistance = 0.5) %>%
#     setConversion(to = "cropland", label = "Intensivierung") %>%
#     setSuitability(const = 1, forest_suit = 1)
#   regLanduse(name = "cropland", resistance = 0.5) %>%
#     setSuitability(const = 1, cropland_suit = 1)
#
#   # register demand types ----
#   regCommodity(name = "crops") %>%
#     setProduction(by = "cropland", amount = 1)
#
#   # register regions ----
#   st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#     mutate(region = "somewhere") %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#          grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#     regDemand(values = "grown", region = "region")
#
#   # check out the landuse change sequence ----
#   showSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   showAnimation(interval = 0.5)
```

## conversion limit

...

```{r}
#| eval: false

model <- "ls2d1_convLimit"
```

```{r}
#| eval: false
#| echo: false

  # clue_gridded(file = paste0(data, "/"),
  #              type = "restrictions")

  # clue_gridded(file = paste0(data, "/suitability2_0.tif"),
  #              type = "conversion", name = "crop_mask") |>
  #              

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing/"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("forest", "cropland")
#   commodities <- c("crops", "trees")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "wuf_masked", start = 2000, end = 2019,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   LUS <- tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1)
#
#   regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
#   setLAT(table = LUS)
#   regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
#   regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest", crs = targetCRS)
#   # make a new object for masking
#   temp <- raster(list.files(path = data_dir, pattern = "suitCropland", full.names = TRUE))
#   temp[temp > 0.7 | temp < 0.5] <- 0
#   temp[temp != 0] <- 1
#   regGridded(name = "feld_masked", obj = temp, crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "forest", resistance = 0.5) %>%
#     setConversion(to = "cropland", mask = "feld_masked", label = "Intensivierung") %>%
#     setSuitability(const = 1, forest_suit = 1)
#   regLanduse(name = "cropland", resistance = 0.5) %>%
#     setSuitability(const = 1, cropland_suit = 1)
#
#   # register demand types ----
#   regCommodity(name = "crops") %>%
#     setProduction(by = "cropland", amount = 1)
#
#   # register regions ----
#   st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#     mutate(region = "somewhere") %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#          grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#     regDemand(values = "grown", region = "region")
#
#   # check out the landuse change sequence ----
#   showSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   showAnimation(interval = 0.5)

```


## protected areas + conversion limit

...

```{r}
#| eval: false

model <- "ls2d1_limit_prot"

```

```{r}
#| eval: false
#| echo: false

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing/"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("forest", "cropland")
#   commodities <- c("crops", "trees")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "wuf_masked_protected", start = 2000, end = 2019,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   LUS <- tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1)
#
#   regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
#   setLAT(table = LUS)
#   regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
#   regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest", crs = targetCRS)
#   # make a new object for masking
#   temp <- raster(list.files(path = data_dir, pattern = "suitCropland", full.names = TRUE))
#   temp[temp > 0.7 | temp < 0.5] <- 0
#   temp[temp != 0] <- 1
#   regGridded(name = "feld_masked", obj = temp, crs = targetCRS)
#   regGridded(name = "region", dir = data_dir, pattern = "protected", crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "forest", resistance = 0.5) %>%
#     setConversion(to = "cropland", mask = "feld_masked", label = "Intensivierung") %>%
#     setSuitability(const = 1, forest_suit = 1)
#   regLanduse(name = "cropland", resistance = 0.5) %>%
#     setSuitability(const = 1, cropland_suit = 1)
#
#   # register demand types ----
#   regCommodity(name = "crops") %>%
#     setProduction(by = "cropland", amount = 1)
#
#   # register regions ----
#   st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#     mutate(region = "somewhere") %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#          grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#     regDemand(values = "grown", region = "region")
#
#   # check out the landuse change sequence ----
#   showSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   showAnimation(interval = 0.5)
```

:::



## minimal example, 

...

```{r}
#| eval: false

model <- "ls2d1_cropRot"

```

```{r}
#| eval: false
#| echo: false

```

## minimal example, dynamic driver

This model has yearly suitability layers (i.e., dynamic drivers).

```{r}
#| eval: false

model <- "ls2d1_dynamic"

```

```{r}
#| eval: false
#| echo: false

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing/"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("forest", "cropland")
#   commodities <- c("crops", "trees")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "2lus_1demand_dynamicDriver", start = 2000, end = 2019,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   LUS <- tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1)
#
#   regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
#   setLAT(table = LUS)
#   regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
#   regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest_year", crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "forest", resistance = 0.5) %>%
#     setConversion(to = "cropland", label = "Intensivierung") %>%
#     setSuitability(const = 1, forest_suit = 1)
#   regLanduse(name = "cropland", resistance = 0.5) %>%
#     setSuitability(const = 1, cropland_suit = 1)
#
#   # register demand types ----
#   regCommodity(name = "crops") %>%
#     setProduction(by = "cropland", amount = 1)
#
#   # register regions ----
#   st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#     mutate(region = "somewhere") %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#          grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#     regDemand(values = "grown", region = "region")
#
#   # check out the landuse change sequence ----
#   showSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   showAnimation(interval = 0.5)

```

## minimal example, location specific preferences

...

```{r}
#| eval: false

model <- "ls2d1_preferences"

```

```{r}
#| eval: false
#| echo: false

# clue_gridded(file = paste0(data, "/"), 
#              type = "preference", name = "")

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing/"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("forest", "cropland")
#   commodities <- c("crops", "trees")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "2lus_1demand_locSpec", start = 2000, end = 2019,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   LUS <- tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1)
#
#   regGridded(name = "landuse", dir = data_dir, pattern = "ls_2lus", crs = targetCRS)
#   setLAT(table = LUS)
#   regGridded(name = "cropland_suit", dir = data_dir, pattern = "suitCropland", crs = targetCRS)
#   regGridded(name = "forest_suit", dir = data_dir, pattern = "suitForest_year0", crs = targetCRS)
#   regGridded(name = "pref_wald", dir = data_dir, pattern = "preferences1", crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "forest", resistance = 0.5) %>%
#     setSuitability(const = 1, forest_suit = 1) %>%
#     setPreference(layer = "pref_wald", weight = 1) %>%
#     setConversion(to = "cropland", label = "Intensivierung")
#   regLanduse(name = "cropland", resistance = 0.5) %>%
#     setSuitability(const = 1, cropland_suit = 1) %>%
#     setPreference()
#
#   # register demand types ----
#   regCommodity(name = "crops") %>%
#     setProduction(by = "cropland", amount = 1)
#
#   # register regions ----
#   st_read(dsn = paste0(data_dir, "example_region.gpkg")) %>%
#     mutate(region = "somewhere") %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   tibble(year = c(2000:2019), region = "somewhere", commodity = "crops",
#          grown = seq(from = 4760, to = 5000, length.out = 20)) %>%
#     regDemand(values = "grown", region = "region")
#
#   # check out the landuse change sequence ----
#   showSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   showAnimation(interval = 0.5)
```



# minimal example + pasture

In a system with cropland, forest and pasture, convert forest to cropland at a rate required only by the demand of crops.

```{r}
#| eval: false

model <- "ls3d1"

minimal <- clue_scene(
  root = paste0(tempdir(), "/", model), 
  name = model, 
  period = c(2000, 2024),
  regions = paste0(data, "/test_region.gpkg"),
  description = "this is a minimal showcase where forest is converted to cropland to produce crops.") |> 
  clue_gridded(file = paste0(data, "/ls3.tif"),
               type = "initial") |> 
  clue_gridded(file = paste0(data, "/suitability1.tif"), 
               type = "driver", 
               name = "cropland_suit") |> 
  clue_gridded(file = paste0(data, "/suitability2_0.tif"),
               type = "driver", 
               name = "forest_suit") |> 
  clue_gridded(file = paste0(data, "/suitability3.tif"),
               type = "driver", 
               name = "pasture_suit") |> 
  clue_landtype(name = "forest", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "forest_suit"),
                                     beta = c(1, 1)),
                conversion = tibble(to = "cropland", 
                                    label = "intensification")) |> 
  clue_landtype(name = "cropland", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "cropland_suit"),
                                     beta = c(1, 1)),
                production = tibble(goods = "crops", 
                                    amount = 1, priority = 1)) |> 
  clue_landtype(name = "pasture", 
                resistance = 0.5,
                suitability = tibble(driver = c("const", "pasture_suit"),
                                     beta = c(1, 1))) |> 
  clue_goods(name = "crops", 
             demand = tibble(year = c(2000:2024), 
                             region = "ravenholm", 
                             amount = seq(from = 5845, to = 6500, length.out = 25))) |> 
  clue_initiate()

clue_run(scene = minimal, clue.exe = ...)
```


# Laos

This is the native CLUMonodo example, here we implement it with the code of `clueRIO`.

```{r}
#| eval: false

model <- "laos"

```

```{r}
#| eval: false
#| echo: false

# also make a usecase where the laos data are pulled from zenodo (need to setup that repo) and then used as conveniently as possible (https://ict.ipbes.net/ipbes-ict-guide/data-and-knowledge-management/technical-guidelines/zenodo)

#   # define metadata ----
#   # dir <- tempdir()
#   # dir <- "/home/se87kuhe/idiv-mount/groups/MAS/01_data/LUCKINet/allocation"
#   dir <- "/media/se87kuhe/external/r-dev/luckiCLU_misc/testing"
#   data_dir <- paste0(system.file("test_datasets", package = "luckiCLU", mustWork = TRUE), "/")
#
#   lus <- c("Rock", "Water", "Urban", "Mining concessions", "Tree plantation concessions",
#            "Arable plantation concessions", "Closed forest", "Shifting cultivation",
#            "Permanent cultivation", "Forest - SC mosaic", "Forest - permanent mosaic")
#   commodities <- c("Built-up", "Staple crop", "Arable cash crop", "Tree cash crop")
#   targetCRS <- "+proj=utm +zone=48 +datum=WGS84 +units=m +no_defs"
#
#   # start project ----
#   startProject(root = dir, name = "laos", start = 2010, end = 2020,
#                landuse = lus, demands = commodities)
#
#   # register gridded objects ----
#   regGridded(name = "landuse", dir = data_dir, pattern = "cov_all.0", crs = targetCRS)
#   tibble(system = lus) %>%
#     mutate(ID = seq_along(system)-1) %>%
#     setLAT()
#
#   # then, set the regular drivers (doesn't need RAT)
#   regGridded(name = "w_3_ab2k", dir = data_dir, pattern = "gr0.fil", crs = targetCRS)
#   regGridded(name = "access", dir = data_dir, pattern = "gr1.fil", crs = targetCRS)
#   regGridded(name = "access_dom", dir = data_dir, pattern = "gr2.fil", crs = targetCRS)
#   regGridded(name = "access_int", dir = data_dir, pattern = "gr3.fil", crs = targetCRS)
#   regGridded(name = "contractp", dir = data_dir, pattern = "gr4.fil", crs = targetCRS)
#   regGridded(name = "popdensity", dir = data_dir, pattern = "gr5.fil", crs = targetCRS)
#   regGridded(name = "slope", dir = data_dir, pattern = "gr6.fil", crs = targetCRS)
#   regGridded(name = "elevation", dir = data_dir, pattern = "gr7.fil", crs = targetCRS)
#   regGridded(name = "precipitation", dir = data_dir, pattern = "gr8.fil", crs = targetCRS)
#   regGridded(name = "temperature", dir = data_dir, pattern = "gr9.fil", crs = targetCRS)
#   regGridded(name = "awc_1", dir = data_dir, pattern = "gr10.fil", crs = targetCRS)
#   regGridded(name = "awc_4", dir = data_dir, pattern = "gr11.fil", crs = targetCRS)
#   regGridded(name = "awc_5", dir = data_dir, pattern = "gr12.fil", crs = targetCRS)
#   regGridded(name = "drain4", dir = data_dir, pattern = "gr13.fil", crs = targetCRS)
#   regGridded(name = "t_clay", dir = data_dir, pattern = "gr14.fil", crs = targetCRS)
#   regGridded(name = "s_clay", dir = data_dir, pattern = "gr15.fil", crs = targetCRS)
#   regGridded(name = "t_gravel", dir = data_dir, pattern = "gr16.fil", crs = targetCRS)
#   regGridded(name = "toc_4", dir = data_dir, pattern = "gr17.fil", crs = targetCRS)
#   regGridded(name = "w_2_un2k", dir = data_dir, pattern = "gr18.fil", crs = targetCRS)
#
#   # register landuse systems ----
#   regLanduse(name = "Rock", resistance = 1)
#   regLanduse(name = "Water", resistance = 1)
#   regLanduse(name = "Urban", resistance = 1)
#   regLanduse(name = "Mining concessions", resistance = 1)
#   regLanduse(name = "Tree plantation concessions", resistance = 0.8) %>%
#     setConversion(to = "Urban", label = "urbanisation")
#   regLanduse(name = "Arable plantation concessions", resistance = 0.6) %>%
#     setConversion(to = "Urban", label = "urbanisation")
#   regLanduse(name = "Closed forest", resistance = 0.2) %>%
#     setConversion(to = "Urban", label = "urbanisation") %>%
#     setConversion(to = "Tree plantation concessions", label = "intensification") %>%
#     setConversion(to = "Arable plantation concessions", label = "intensification") %>%
#     setConversion(to = "Shifting cultivation", label = "intensification") %>%
#     setConversion(to = "Permanent cultivation", label = "intensification") %>%
#     setConversion(to = "Forest - SC mosaic", label = "intensification") %>%
#     setConversion(to = "Forest - permanent mosaic", label = "intensification")
#   regLanduse(name = "Shifting cultivation", resistance = 0.2) %>%
#     setConversion(to = "Urban", label = "urbanisation") %>%
#     setConversion(to = "Tree plantation concessions", label = "intensification") %>%
#     setConversion(to = "Arable plantation concessions", label = "intensification") %>%
#     setConversion(to = "Closed forest", label = "abandonment")
#   regLanduse(name = "Permanent cultivation", resistance = 0.45) %>%
#     setConversion(to = "Urban", label = "urbanisation") %>%
#     setConversion(to = "Tree plantation concessions", label = "intensification") %>%
#     setConversion(to = "Arable plantation concessions", label = "intensification") %>%
#     setConversion(to = "Closed forest", label = "extensification") %>%
#     setConversion(to = "Forest - permanent mosaic", label = "extensification")
#   regLanduse(name = "Forest - SC mosaic", resistance = 0.2) %>%
#     setConversion(to = "Urban", label = "urbanisation") %>%
#     setConversion(to = "Tree plantation concessions", label = "intensification") %>%
#     setConversion(to = "Arable plantation concessions", label = "intensification") %>%
#     setConversion(to = "Closed forest", label = "extensification") %>%
#     setConversion(to = "Shifting cultivation", label = "expansion")
#   regLanduse(name = "Forest - permanent mosaic", resistance = 0.2) %>%
#     setConversion(to = "Urban", label = "urbanisation") %>%
#     setConversion(to = "Tree plantation concessions", label = "intensification") %>%
#     setConversion(to = "Arable plantation concessions", label = "intensification") %>%
#     setConversion(to = "Closed forest", label = "extensification") %>%
#     setConversion(to = "Permanent cultivation", label = "intensification")
#
#   # set suitability ----
#   read_csv(file = paste0(data_dir, "params.csv")) %>%
#     setSuitability(table = .)
#
#   # register demand types ----
#   regCommodity(name = "Built-up") %>%
#     setProduction(by = "Urban", amount = 400)
#   regCommodity(name = "Staple crop") %>%
#     setProduction(by = "Permanent cultivation", amount = 2.72) %>%
#     setProduction(by = "Shifting cultivation", amount = 2.08) %>%
#     setProduction(by = c("Forest - SC mosaic", "Forest - permanent mosaic"), amount = c(0.80, 1.44))
#   regCommodity(name = "Arable cash crop") %>%
#     setProduction(by = "Arable plantation concessions", amount = 6.40)
#   regCommodity(name = "Tree cash crop") %>%
#     setProduction(by = "Tree plantation concessions", amount = 6.40)
#
#   # register regions ----
#   read_rds(path = paste0(data_dir, "gadm36_LAO_0_sf.rds")) %>%
#     mutate(region = tolower(NAME_0)) %>%
#     dplyr::select(region, geometry, -GID_0, -NAME_0) %>%
#     regRegions(region = "region")
#
#   # register demand ----
#   read_csv(file = paste0(data_dir, "demands.csv")) %>%
#     dplyr::select(year, commodity = type, region, values = demand) %>%
#     regDemand(values = "values", region = "region")
#
#   # check out the landuse change sequence ----
#   makeSequence()
#
#   # prepare input files ----
#   makeModel()
#
#   # and finally run the model ----
#   runModel()
#
#   # make an animation ----
#   makeAnimation()

```


```{r}
#| eval: false
#| echo: false

# alternative way that may be better suited to batch processing
# .suit <- tibble(type = c("forest", "forest", "cropland", "cropland"),
#                 driver = c("const", "forest_suit", "const", "cropland_suit"),
#                 beta = c(1, 1, 1, 1))
# .conv <- tibble(type = "forest",
#                 to = "cropland", 
#                 label = "intensification")
# .prod <- tibble(goods = "crops", 
#                 amount = 1,
#                 priority = 1)
# 
# minimal <- minimal |> 
#   clue_landtype(name = "forest", resistance = 0.5,
#                 suitability = .suit |> filter(type == "forest"),
#                 conversion = .conv |> filter(type = "forest")) |> 
#   clue_landtype(name = "cropland", resistance = 0.5,
#                 suitability = .suit |> filter(type == "cropland"),
#                 production = .prod |> filter(goods == "crops"))

```

# template 

...

```{r}
#| eval: false

model <- ""

```

```{r}
#| eval: false
#| echo: false

```


# Outro

If you have any questions, issue reports or (kind) comments, please leave them below!
